SubGenome Analysis

### 2017-07-27 R/SubGenome Analysis.Rmd
Report table of matching chromosomal segments
Also, warn if there are unmatched chromosomes.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readr)
library(plyr)
library(reshape2)
library(ggplot2)
library(readxl)
```
```{r preamble, echo=FALSE}
paste("This analysis is using the file", params$inputDataFile)
```

@@

> manually perform the parse, correct the header lines, and run the import lines until pipeline is debugged
> perl parseKsKnFile.pl ../Data/6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum_v1_vs_maize_v1.tab

```{r importSynMapData, include=FALSE}
cmd <- paste("perl", params$fileParser, params$synMapOutputFile)
system(cmd)
syntelogs.import <- read_delim(params$inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]
```
```{r reportSynMapData, echo=FALSE}
# Report results of import
sprintf("We have imported %d lines from the parsed SynMap file and removed %d incomplete rows resulting in %d syntenic genes.", nrow(syntelogs.import), nrow(syntelogs.import)- nrow(syntelogs.complete), nrow(syntelogs.complete))
```

@@

```{r findSubgenomes, echo=FALSE}
## Find which sets of chromosomes with syntelogs should be in group 1 or group 2, where group 1 has larger syntenic blocks
# First group rows with same syntenic block and chromosome, then summarize each block's ks values with median/mean/count
syntelogs.aggregate <- ddply(syntelogs.complete, ~block+org_chr1+org_chr2, summarise, median=median(ks), mean=mean(ks), count=length(ks))

# Following the schnable article, syntenic blocks must have 12 genes "The median synonymous substitution rate of all gene pairs in a syntenic block between maize and sorghum can be used to classify syntenic blocks of 12 or more genes unambiguously as orthologous or homoeologous, however" and have a median ks value that discriminates for the alpha duplication event.
syntelogs.orthologs <- subset(syntelogs.aggregate, count>=12 & median<=ks_cutoff)

# Determine which chromosome has largest syntenic block, the syntelogs on this chromosome are subgenome 1
orthologs.counts <- ddply(syntelogs.orthologs, ~org_chr1+org_chr2, summarise, total_genes=sum(count))
orthologs.sorted <- orthologs.counts[order(orthologs.counts["org_chr1"], -orthologs.counts["total_genes"]),]

# orthologs.sorted[match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1),] #first instance of each group (i.e. biggest=sub1)
orthologs.sub1rows <- match(unique(orthologs.sorted$org_chr1), orthologs.sorted$org_chr1) #not exactly right!!!! how to split sub 1 over multiple chromosomes?
orthologs.sub1 <- orthologs.sorted[orthologs.sub1rows,]
orthologs.sub2 <- orthologs.sorted[-orthologs.sub1rows,]
orthologs.sub1$subGenome="sub1"
orthologs.sub2$subGenome="sub2"
```

@@

```{r writeSubgenomesToFile, echo=FALSE}
## Now recover which genes are in these sets
subgenome.groups <- rbind(merge(syntelogs.orthologs, orthologs.sub1, all.x = F), merge(syntelogs.orthologs, orthologs.sub2, all.x = F))
subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
write.table(subgenome, params$outputFile, sep="\t")
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(plyr)#
library(dplyr)
library(tidyr)
library(reshape2)#
library(ggplot2)
```
```{r preamble, echo=FALSE}
paste("This analysis is using the file", params$inputDataFile)
```

@@

> manually perform the parse and run the import lines until pipeline is debugged
> perl parseKsKnFile.pl ../Data/6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum_v1_vs_maize_v1.tab

```{r importSynMapData, include=FALSE}
cmd <- paste("perl", params$fileParser, params$synMapOutputFile)
system(cmd)
syntelogs.raw <- read_delim(params$inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
syntelogs.complete <- syntelogs.raw[complete.cases(syntelogs.raw),]
```
```{r reportSynMapData, echo=FALSE, results="asis"}
# Report results of import
cat(sprintf("We have imported %d lines from the parsed SynMap file and removed %d incomplete rows resulting in %d syntenic genes.", nrow(syntelogs.raw), nrow(syntelogs.raw)- nrow(syntelogs.complete), nrow(syntelogs.complete)))
```

@@

```{r findSubgenomes, include=FALSE}
## Find which sets of chromosomes with syntelogs should be in group 1 or group 2, where group 1 has larger syntenic blocks
# First group rows with same syntenic block and chromosome, then summarize each block's ks values with median/mean/count
syntelogs.aggregate <- ddply(syntelogs.complete, ~block+org_chr1+org_chr2, summarise, median=median(ks), mean=mean(ks), count=length(ks))

# Following the schnable article, syntenic blocks must have 12 genes "The median synonymous substitution rate of all gene pairs in a syntenic block between maize and sorghum can be used to classify syntenic blocks of 12 or more genes unambiguously as orthologous or homoeologous, however" and have a median ks value that discriminates for the alpha duplication event.
homeologs <- subset(syntelogs.aggregate, count>=12 & median<=ks_cutoff)

# Determine which chromosome has largest syntenic block, the syntelogs on this chromosome are subgenome 1
homeologs.counts <- ddply(homeologs, ~org_chr1+org_chr2, summarise, total_genes=sum(count))
homeologs.sorted <- homeologs.counts[order(homeologs.counts["org_chr1"], -homeologs.counts["total_genes"]),]

# homeologs.sorted[match(unique(homeologs.sorted$org_chr1), homeologs.sorted$org_chr1),] #first instance of each group (i.e. biggest=sub1)
homeologs.sub1rows <- match(unique(homeologs.sorted$org_chr1), homeologs.sorted$org_chr1) #not exactly right!!!! how to split sub 1 over multiple chromosomes?
homeologs.sub1 <- homeologs.sorted[homeologs.sub1rows,]
homeologs.sub2 <- homeologs.sorted[-homeologs.sub1rows,]
homeologs.sub1$subGenome="sub1"
homeologs.sub2$subGenome="sub2"

# Reconstructed ancestral chromosomes
homeologs %>% group_by(org_chr1, org_chr2) %>% summarize(genes=sum(count)) %>% spread(org_chr2, genes) -> homeologs.spread
homeologs.spread$total <- rowSums(homeologs.spread[-1], na.rm = TRUE)
syntelogs.complete %>% select(org_chr1) %>% distinct() %>% left_join(homeologs.spread) %>% replace(is.na(.), 0) %>% arrange(desc(total)) -> homeologs.totals
```
```{r reportFindSubgenomes, echo=FALSE, results="asis"}
homeologs.totals %>% knitr::kable(caption = "Table")
homeologs.totals %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0) %>% sprintf("There are %i unaligned sorghum chromosome(s)", .) %>% cat()

```

@@

```{r writeSubgenomesToFile, echo=FALSE}
## Now recover which genes are in these sets
subgenome.groups <- rbind(merge(homeologs, homeologs.sub1, all.x = F), merge(homeologs, homeologs.sub2, all.x = F))
subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
write.table(subgenome, params$outputFile, sep="\t")
```
